---
title: "Lab 3"
author: "Brady Lamson"
format: pdf
editor: visual
---

```{r}
data <- read.table(header = T, file = "../datasets/properties.txt")
```


# 1.1

```{r, echo=FALSE, include=FALSE}
plot(y= data$Rent.Rate, x=data$Op.Expense, xlim = c(0,15), ylim = c(0,20), pch=20)
```


## 3. 

```{r}
data_reg <- lm(Rent.Rate ~ -1 + Op.Expense, data = data)
data_reg |> summary()
```

$$
Y_i = 1.478X_i + \epsilon_i
$$

## 4. 

Forcing the line to go through the origin does **not** seem appropriate for this data set.

## 5. 

The pattern seems relatively flat, maybe a slight positive relationship as theres a larger clustering around the top on the righthand side than the left.

## 6. 

The residuals do not sum to zero, they sum to 66.27

# Part B

## 2.

```{r}
standard_reg <- lm(Rent.Rate ~ Op.Expense, data = data)
X <- model.matrix(standard_reg)

solve(t(X) %*% X)
```

## 4.

```{r}
Y <- data$Rent.Rate
b <- solve(t(X) %*% X) %*% (t(X) %*% Y)
b
```

The coefficients in $\vec{b}$ are the same as the output of summary.

## 5.

```{r}
y_hat <- X %*% b
y_hat |> head()
```

These values are the same as the fitted values from the regression.

## 6.

```{r}
(Y - y_hat) |> head()
```

These match the residuals of the regression.

## 7. 

```{r, echo=FALSE, include=FALSE}
standard_reg |> summary()
```

```{r}
sqrt(1.575^2 * solve(t(X) %*% X))
```

The diagonals are the same as the std. error from the output of summary.

# Part C

```{r}
length <- c(63, 63, 63, 63, 63, 63, 63, 63, 63)
weight<- c(136, 198, 194, 140, 93, 172, 116, 174, 145)
snakes <- data.frame(Length = length, Weight = weight)

my.reg <- lm(Weight ~ Length, data = snakes)
```

There's only 1 unique value for length, so that ruins everything here. 

