---
title: "Homework 4"
author: "Brady Lamson"
format: pdf
editor: visual
---

```{r}
gpa <- read.table("../datasets/gpa.txt", header = T)
```

# Problem 4.14
## Part A
```{r}
no_int_lm <- lm(gpa ~ 0 + act.score, data = gpa)
no_int_lm |> summary()
```

$$Y = 0 - 3.0276X_1 + \epsilon$$

## Part B

```{r}
conf <- confint(no_int_lm, level = .95)
glue::glue(
    "With 95% confidence, the slope of beta_1 is in  
    ({conf[1] |> round(3)}, {conf[2] |> round(3)})
    "
)
```

As the confidence interval does not contain 0, there is significant evidence to indicate that, in the no intercept model, there is a positive relationship between act scores and college freshman gpa.

## Part C

```{r}
my.new.data <- data.frame(act.score = 30)
predict(no_int_lm, newdata = my.new.data, interval = "confidence",
level = 0.95)
```

# Problem 4.15

## Part A

```{r}
plot(
    x = gpa$act.score, y = gpa$gpa, pch=20, 
    xlim = c(0,35), ylim = c(0,4)
)
abline(no_int_lm)

```

The linear regression though the origin does not appear to be a good fit. 

## Part B

```{r}
sum_resid <- no_int_lm$residuals |> sum()

glue::glue("The sum of the residuals is {sum_resid |> round(3)}")
```

The residuals do not sum to 0.

```{r}
plot(no_int_lm$fitted.values, no_int_lm$residuals, pch=20)
```

The residuals seem to trend downward as the fitted values increase. 

## Part C

```{r}
full_reg <- lm(gpa ~ act.score, data = gpa)

anova(no_int_lm, full_reg)
```

$$
\begin{aligned}
H_0: E(Y) &= \beta_0 + \beta_1 X \\
H_a: E(Y) &\neq \beta_0 + \beta_1 X \\
\alpha &= 0.005 \\ 
F &= 43.401 \\
p &\approx 1.304 \cdot 10^{-9}
\end{aligned}
$$

As $p < \alpha$, there is significant evidence to indicate the linear regression model is not sufficient to explain the relationship in the data. 

# Problem 5.5

```{r}
finance <- read.table("../datasets/CH05PR05.txt", header = F)
```

## Part 1

```{r}
t(finance$V2) %*% finance$V2
```

## Part 2

```{r}
t(finance$V1) %*% finance$V1
```

```{r}
t(finance$V1) %*% finance$V2
```

# Problem 5.13

```{r}
solve(t(finance$V1) %*% finance$V1)
```

# Problem 5.24

