---
title: "Homework 3"
author: "Brady Lamson"
format: pdf
editor: visual
---

```{r}
# setup datasets

ch1_27 <- read.table("CH01PR27.txt")
colnames(ch1_27) <- c("muscle_mass", "age")
ch1_28 <- read.table("CH01PR28.txt")
colnames(ch1_28) <- c("crime", "diploma_percent")
ch3_15 <- read.table("CH03PR15.txt")
colnames(ch3_15) <- c("concentration", "time")
```


# 2.29

## a

```{r}
muscle_reg <- lm(muscle_mass ~ age, data = ch1_27)
muscle_resids <- muscle_reg$residuals
muscle_deltas <- muscle_reg$fitted.values - mean(ch1_27$muscle_mass)
```

```{r}
plot(x = ch1_27$age, y = muscle_resids, pch=20, ylim = c(-25,25))
```

```{r}
plot(x = ch1_27$age, y = muscle_deltas, pch=20, ylim = c(-25,25))
```

SSR appears to be a larger part of SSTO, which means $r^2$ should be pretty high as a lot of the variance is explained by the regression.

## b

```{r}
muscle_reg |> anova()
```

## c

$$
\begin{aligned}
H_0: \beta_1 = 0\\
H_a: \beta_1 \neq 0\\
\alpha = 0.05\\
F = 174\\
p=2.2 \cdot 10^{-16}
\end{aligned}
$$

As $p < \alpha$ we have sufficient evidence to reject the null hypothesis that the age and muscle mass have no relationship.

## d

.25

## e

$$
\begin{aligned}
r^2 = .75 \\
r = -.87
\end{aligned}
$$

# 2.30

```{r}
crime_reg <- lm(crime ~ diploma_percent, data = ch1_28)
crime_reg |> summary()
```

$$
\begin{aligned}
H_0: \beta_1 = 0\\
H_a: \beta_1 \neq 0\\
\alpha = 0.01\\
t = -4.103\\
p=9.57 \cdot 10^{-5}
\end{aligned}
$$

as $p < \alpha$ there is sufficient evidence to reject the null hypothesis that there is no relationship between crime rates and diploma percentages for an area.

## b

```{r}
confint(crime_reg, level=.99)
```

The slope of the line indicating how crime rate changes as diploma percentage increases is anywhere between -280 and -61 with a 99% confidence.

# 2.31

## a

```{r}
crime_reg |> anova()
```

## b

$$
\begin{aligned}
H_0: \beta_1 = 0\\
H_a: \beta_1 \neq 0\\
\alpha = 0.01\\
F = 16.834\\
p=9.57 \cdot 10^{-5}
\end{aligned}
$$

the p-value is the same as the previous test and so is the conclusion. 

## c

17% of the variation in crime rate is explained by the relationship between it and diploma rates. This is actually pretty high when you think about how many things may influence crime!

## d

$$
\begin{aligned}
r^2 = .1703 \\
r = -.41
\end{aligned}
$$


# 2.32

## a

$$
\begin{aligned}
Y_i = 2015 - 170X_i + \epsilon_i \\
Y_i = 7111 + \epsilon_i
\end{aligned}
$$

## b

```{r}
crime_reg_r <- lm(crime ~ 1, data = ch1_28)
anova(crime_reg_r, crime_reg)
```


$$
\begin{aligned}
SSE_r&= 548736108\\
SSE_f&= 455273165\\
df_r &= 82\\
df_f&= 83\\
F^* &=16.834\\
decision &= reject\; H_0
\end{aligned}
$$

## c

The test statistic and decision rule are equivalent to the earlier results, yes.

# 3.8 (skip part e)

## a

```{r}
stem(ch1_28$diploma_percent)
```

This plot indicates that most counties have between 75% and 84% of their population having diplomas.

## b

```{r}
boxplot(crime_reg$residuals)
```

The residuals appear to be symmetrical, yes.

## c

```{r}
plot(crime_reg$residuals, crime_reg$fitted.values)
```

This tells us that regardless of the fitted values the residuals tend to vary about the same amount.

## d

```{r}
my.pts <- qqnorm(crime_reg$residuals, pch=20)
my.pts
qqline(crime_reg$residuals, col = "blue", lwd=3)
```

```{r}
cor(my.pts$x, my.pts$y)
```

Based on the plot and the result above this appears to pass the normality assumption.

# 3.15**

## a

```{r}
chem_reg <- lm(concentration ~ time, data = ch3_15)
chem_reg |> summary()
```

## b

```{r}
chem_reg_f <- lm(concentration ~ factor(time), data = ch3_15)
chem_reg_r <- lm(concentration ~ 1, data = ch3_15)

anova(chem_reg_r, chem_reg_f)
```

As the p-value is far below the alpha we can reject the null hypothesis that this relationship is linear. 

## c

I don't think the rest really indicates what would work better.

# 3.16 (skip part b)

## a

plot(y=ch3_15$concentration, x=ch3_15$time)

I'll probably do a log transformation, mostly because part c hints at it being the right way to go.

## c


```{r}
chem_trans_reg <- lm(log10(concentration) ~ time, data = ch3_15)
chem_trans_reg |> summary()
```

$$
Y'_i = .654 - .195X_i + \epsilon
$$

## d

```{r}
x <- c(1:9)
y <- .654 - (.195*x)

plot(x=ch3_15$time, y=log10(ch3_15$concentration), pch=20)
lines(x,y, type = 'l', col = "blue", lwd=2)
```

## e

```{r}
plot(x=chem_trans_reg$fitted.values, y=chem_trans_reg$residuals, pch=20)

```

```{r}
qqnorm(chem_trans_reg$residuals, pch=20)
qqline(chem_trans_reg$residuals, col = "blue", lwd=2)
```

What we see is the variance appears constant and the residuals appear normal. So the transformation was effective!